\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,hyperref,array,xcolor,multicol,verbatim,mathpazo}
\usepackage[normalem]{ulem}
\usepackage[pdftex]{graphicx}
\usepackage{fullpage}

\usepackage{threeparttable}
\usepackage{geometry}
\usepackage[format=hang,font=normalsize,labelfont=bf]{caption}
\usepackage{lscape}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{float,color}
\usepackage[pdftex]{graphicx}
\usepackage{pdfsync}
\usepackage{placeins}
\usepackage{geometry}
\usepackage{pdflscape}
\usepackage[normalem]{ulem}
\usepackage{threeparttable, multirow}
\useunder{\uline}{\ul}{}
\synctex=1
\usepackage{hyperref}
\hypersetup{colorlinks,linkcolor=red,urlcolor=blue,citecolor=blue}
\usepackage{bm}


\newcommand{\E}{\mathbb{E}}

% Identifying information
\title{
Homework 3
} 
\author{Linghui Wu}
\date{\today}

\begin{document}

\maketitle

\textbf{(a)} The firm is on the labor demand curve because the real wage is equal to the marginal productivity of labor 
\begin{equation}
\label{eq:labor_demand}
\frac{W_{1}}{P_{1}} = \frac{W_{0}}{P_{1}} = A_{1} = MPL_{1}.
\end{equation}

\textbf{(b)} The household is not on the labor demand curve because the labor-leisure condition does not hold in the short run when the wage is sticky
\begin{equation}
\label{eq:labor_supply}
\frac{W_{1}}{P_{1}} = \frac{W_{0}}{P_{1}} \neq \frac{\chi N^{\varphi}_{1}}{C^{-\gamma}_{1}}.
\end{equation}

\textbf{(c)} At $t=1$ with $W_{0}=W_{1}$, the firm adjusts labor demand $N^{d}$ s.t. equation (\ref{eq:labor_demand}) holds. 
The household is not on the labor supply curve and works as many hours as needed to produce the required output $Y_{1}$.

\textbf{(d)} In the long-run ($t \geq 2$), we have the following characterizations,
\begin{align*}
Y &= AN \\
\frac{W}{P} &= A \\
\frac{W}{P} &= \frac{\chi N^{\varphi}}{C^{-\gamma}} \\
Y &= C \\
\frac{M}{P} &= \zeta^{\frac{1}{\nu}}(1-\frac{1}{Q})^{-\frac{1}{\nu}}C^{\frac{\gamma}{\nu}} \\
1 &= \beta \frac{Q}{\pi}
\end{align*}
where $\pi$ denotes the steady-state inflation rate.
We can then solve for steady state,
\begin{align*}
C &= \left(\frac{A^{\varphi+1}}{\chi}\right)^{\frac{1}{\varphi+\gamma}} \\
Y &= C \\
\frac{W}{P} &= A \\
N &= \left(\frac{A^{1-\gamma}}{\chi}\right)^{\frac{1}{\varphi+\gamma}} \\
Q &= \frac{\pi}{\beta} \\
\frac{M}{P} &= \zeta^{\frac{1}{\nu}}(1-\frac{\beta}{\pi})^{-\frac{1}{\nu}}C^{\frac{\gamma}{\nu}}
\end{align*}

\textbf{(e)} By part (d), we can see that the Classical Dichotomy holds in the long run because the real variables $Y$, $C$, $N$ and $\frac{W}{P}$ are independent of money supply $M$ and nominal interest rate $Q$.

\textbf{(f)} In the short-run, we have 
\begin{align*}
Y_{1} &= A_{1}N_{1} \\
W_{1} &= W_{0} \\
\frac{W_{1}}{P_{1}} &= A_{1} \\
Y_{1} &= C_{1} \\
\frac{M_{1}}{P_{1}} &= \zeta^{\frac{1}{\nu}}\left(1-\frac{1}{Q_{1}}\right)^{-\frac{1}{\nu}}C^{\frac{\gamma}{\nu}}_{1} \\
1 &= \beta Q_{1} \frac{P_{1}}{P} \left(\frac{C}{C_{1}}\right)^{-\gamma}
\end{align*}
where $P$ and $C$ are steady-state price levels and consumption.

Plugging $P_{1}=\frac{W_{1}}{A_{1}}$, the following system of two equations pins down $C_{1}$ and $\frac{M_{1}}{P_{1}}$ which are required for the output and money markets clearing.
\begin{align}
\label{eq:euler_eq}
C_{1} &= Y_{1} = \left(\frac{1}{\beta Q_{1}}\right)^{\frac{1}{\gamma}}C \\
\label{eq:money_demand}
\frac{M_{1}}{P_{1}} &= \zeta^{\frac{1}{\nu}}\left(1-\frac{1}{Q_{1}}\right)^{-\frac{1}{\nu}}C^{\frac{\gamma}{\nu}}_{1} 
\end{align}

\textbf{(g)} Part (f) suggests that the Classical Dichotomy does not hold in the short-run because the sticky wage assumption replaces the household's labor supply curve and short-term consumption is affected by nominal interest rate.

\textbf{(h)} When money supply $M_{1}$ increases, the price of money $Q_{1}$ falls. 
Equation (\ref{eq:euler_eq}) suggests that real interest rate drops, 
and the household then increases today's consumption $C_{1}$ relative to future consumption through intertemporal substitution. 
So there is an overall increase in consumption today, 
which is accommodated by the household working extra hours to produce more output.

\textbf{(i)} In the long run, an increase in productivity leads to increases in 
real wage according to part (d).

In the short-run where the wage is sticky, higher productivity results in an increase in the marginal productivity of labor, and the firm will demand less labor $N^{d}$. 
Changes in $A_{1}$ do not affect the real output and money market.  

\textbf{(j)} The labor wedge at $t=1$ is
\begin{align*}
1-\tau^{N}_{1} &= \frac{MRS_{1}}{MPL_{1}} = \chi A^{\gamma-1}_{1}N^{\gamma+\varphi}_{1}
\end{align*}
In a recession, $N_{1}$ is low and $\tau^{N}_{1}$ is high.
So the labor wedge is countercyclical in the sticky wage model.

\textbf{(k)} We can use data on unemployment and labor share of income - as a proxy for marginal productivity of labor - to distinguish the predictions of sticky price and sticky wage model.

\end{document}


